\documentclass[aspectratio=1610]{beamer}
\usepackage{iftex}
\usepackage{pdfpages}
\ifLuaTeX\else
	\usepackage[utf8]{inputenc}
	\usepackage[T1]{fontenc}
\fi
\usetheme[
% titleimagecolor=red,       % [gray], darkgray, red, blue, green
% titleimagemargin=2mm,      % Distance [2mm]    Frame around title page image
% navigationsymbols=false,   % true   / [false]  Navigation symbols in the foot
% mathseriffont=false,       % true   / [false]  Serif / non-serif math fonts
% foot=true,                 % [true] / false    Footline or not
% nofootslidenum=false       % true   / [false]  Keep slide num even when foot=false
% footlogo=true,             % [true] / false    Put LU logo to the left of footer
% english=true,              % [true] / false    English / Swedish logo
% LTHlogo=false,             % true   / [false]  Use LTH logo instead of LU on title and end pages.
% blackenumeratenumber=true, % [true] / false    Black enumerate numbers, o.w. Lund bronze
% blackitemmark=false,       % true   / [false]  Black item marks, o.w. Lund bronze
% defaultfont=palatino,      % [palatino], beamer, lu
% sectionframe=true,
]{ulund}
%%%%%%%%%%%%%%%%%%%%% Layout commands 
%%%% Foot
% \ulundfootleft{\insertshortauthor}
% \ulundfootmid{\insertshorttitle}
% \ulundfootright{\insertframenumber}% {\insertframenumber:\inserttotalframenumber}
%%%% Titleimage
% \titleimage{Pictures/ULUNDcolor} % Replaces the LU image. Voids option titleimagecolor
%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%

\newcommand{\code}[1]{\lstinline{#1}}
\title[JavaScript]{JavaScript}
\author[EDAF90]{%
  EDAF90 Web Programming\newline
  Per Andersson}
%%%%%%%%%%%%%%%%%%%%%
%\usepackage{verbatim}
%%%%%%%%%%%%% Verbatim code box
\usepackage[skins,listings]{tcolorbox}
\ifLuaTeX\else
	\tcbuselibrary{listingsutf8}
\fi
\newtcblisting{CodeBox}[2][]{% Only code
  colframe=black,
  colback=white,
  arc=1pt,
  boxrule=0.5pt,
  top=0mm,bottom=0pt,left=0pt,
  colbacktitle=gray!40,
  coltitle=black,
  fonttitle=\sffamily,
  listing only,
  hbox,
  listing options={
    basicstyle=\small\ttfamily,
    breaklines=true,
    columns=fullflexible,
    language=Java,
    basicstyle = \ttfamily,
    showstringspaces=false,
  },
  title=#2,#1
}

\lstset{
    basicstyle=\small\ttfamily,
    breaklines=true,
    columns=fullflexible,
    language=Java,
    basicstyle = \ttfamily,
    showstringspaces=false,
    morekeywords= {undefined, NaN, function, export, var, let, of, in}
}

%%%%%%%%%%%%%%%%%%%%%
%%%%%%%%%%%%%%%%%%%%%
%%%%%%%%%%%%%%%%%%%%%

\begin{document}
\begin{frame}[plain]% Use plain to suppress footline box
  \titlepage
\end{frame}

%%%%%%%%%%%%%%%
\begin{frame}
  \frametitle{JavaScript}

\begin{itemize}
\item ``the world's most misunderstood programming language''
\item working name at Netscape 1995: \emph{LiveScript}
\item syntax and name in common with Java
\item object function oriented language
\item interpreted
\item dynamically typed 
\item run in any web browser and node.js
\end{itemize}
\end{frame}

%%%%%%%%%%%%%%%
\begin{frame}
  \frametitle{Interpreted}

\begin{itemize}
\item no compilation $\rightarrow$ no compilation errors
\item design decision: try to execute everything
  \begin{itemize}
  \item many silent errors
  \item weird and unexpected behaviour
  \item for example a miss spelled  property name
  \end{itemize}
\item the programmer have more responsibility
\item programmers needs extended language knowledge
\item use jslint to check your code
\item write test cases to catch compilation errors
\end{itemize}
\end{frame}


%---Types ---------------------------------------------------
\include{parts/types}

%---Truthy/Falsy ---------------------------------------------------
\include{parts/truthy_equality}

%---Functions ---------------------------------------------------
\include{parts/functions}

%---scope ---------------------------------------------------
\include{parts/scope}

%---modules ---------------------------------------------------
\include{parts/modules}

%---objects ---------------------------------------------------
\include{parts/objects}

%---arrays ---------------------------------------------------
\include{parts/arrays}

%---Destructuring assignment ---------------------------------------------------
\include{parts/destructuring_assignment}

%--- Automatic Semicolon Insertion ---------------------------------------------------
\begin{frame}[fragile]
  \frametitle{Automatic Semicolon Insertion}
  If needed, a semicolon is added at the end of a line.
\begin{CodeBox}{the function will return undefined}
  function() { return
  1; }
\end{CodeBox}

\vspace{4mm}
Common to use minify to minimise script download size. All white spaces are removed.
\begin{columns}[onlytextwidth]
  \begin{column}{0.35\textwidth}
\begin{CodeBox}{works}
  let myVar = 9
  if (myVar === 9) {
  }
\end{CodeBox}
  \end{column}
  \begin{column}{0.65\textwidth}
\begin{CodeBox}{syntax error after minify}
  var myVar = 9 if (myVar === 9) {}
\end{CodeBox}
  \end{column}
  \begin{column}{0.3\textwidth}  \end{column}
\end{columns}%

\vspace{4mm}
Use jslint to detect these problems.
\end{frame}

%---Strict mode ---------------------------------------------------
\begin{frame}[fragile]
  \frametitle{Strict mode}
Converting mistakes into errors.

\begin{CodeBox}{Whole-script strict mode syntax}
'use strict';
var v = "Hi! I'm a strict mode script!";
\end{CodeBox}

\begin{CodeBox}{Function-level strict mode syntax}
function strict() {
  'use strict';
  function nested() { return 'And so am I!'; }
  return "Hi!  I'm a strict mode function!  " + nested();
}
function notStrict() { return "I'm not strict."; }
\end{CodeBox}


\end{frame}


%%%%%%%%%%%
\end{document}
