\documentclass[aspectratio=1610]{beamer}
\usepackage{iftex}
\usepackage{pdfpages}
\ifLuaTeX\else
	\usepackage[utf8]{inputenc}
	\usepackage[T1]{fontenc}
\fi
%%%%%%%
% \usepackage{layout}
% \usepackage{lipsum}
%%%%%%%
\usetheme[% Complete settings. Default value in []
% titleimagecolor=red,       % [gray], darkgray, red, blue, green
% titleimagemargin=2mm,      % Distance [2mm]    Frame around title page image
% navigationsymbols=false,   % true   / [false]  Navigation symbols in the foot
% mathseriffont=false,       % true   / [false]  Serif / non-serif math fonts
% foot=true,                 % [true] / false    Footline or not
% nofootslidenum=false       % true   / [false]  Keep slide num even when foot=false
% footlogo=true,             % [true] / false    Put LU logo to the left of footer
% english=true,              % [true] / false    English / Swedish logo
% LTHlogo=false,             % true   / [false]  Use LTH logo instead of LU on title and end pages.
% blackenumeratenumber=true, % [true] / false    Black enumerate numbers, o.w. Lund bronze
% blackitemmark=false,       % true   / [false]  Black item marks, o.w. Lund bronze
% defaultfont=palatino,      % [palatino], beamer, lu
% sectionframe=true,
]{ulund}
%%%%%%%%%%%%%%%%%%%%% Layout commands 
%%%% Foot
% \ulundfootleft{\insertshortauthor}
% \ulundfootmid{\insertshorttitle}
% \ulundfootright{\insertframenumber}% {\insertframenumber:\inserttotalframenumber}
%%%% Titleimage
% \titleimage{Pictures/ULUNDcolor} % Replaces the LU image. Voids option titleimagecolor
%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%
\title[EDAF90 Web Programming]{EDAF90 Web Programming}
\author[Per Andersson]{%
  Per Andersson\newline
  Dept.\@ of CS, Lund University}
%%%%%%%%%%%%%%%%%%%%%
%\usepackage{verbatim}
%%%%%%%%%%%%% Verbatim code box
\usepackage[skins,listings]{tcolorbox}
\ifLuaTeX\else
	\tcbuselibrary{listingsutf8}
\fi
\lstdefinestyle{CodeStyle}{basicstyle = \ttfamily}
\newtcblisting{CodeBox}[2][]{% Only code
  colframe=black,
  colback=white,
  arc=1pt,
  boxrule=0.5pt,
  top=0mm,bottom=0pt,left=0pt,
  colbacktitle=gray!40,
  coltitle=black,
  fonttitle=\sffamily,
  listing only,
  title=#2,#1}
%%%%%%%%%%%%%%%%%%%%%
%%%%%%%%%%%%%%%%%%%%%
%%%%%%%%%%%%%%%%%%%%%
\begin{document}
\begin{frame}[plain]% Use plain to suppress footline box
  \titlepage
\end{frame}

%%%%%%%%%%%%%%%
\begin{frame}
  \frametitle{Course Content}

This course is about programming apps running in a web browser.

\vspace{5mm}

During the course you will:
\begin{itemize}
\item learn web terminology and consepts
\item learn basic principles for HTLM and CSS
\item build applications for web browsers
\item understand some internet protokolls
\item learn basic asynchronous programming and streams
\item be introduced to concurrency, race conditions, and deadlocks
\end{itemize}

\end{frame}


%%%%%%%%%%%%%%%
\begin{frame}
  \frametitle{Not Covered in the Course}
I assume you know class based object oriented programming.
\\ \vspace{5mm}
The course do not cover:
\begin{itemize}
\item advanced or a complete guide to HTML
\item advanced or a complete guide to  CSS
\item GUI design/interaction design
\item security
\item backend
\item software architectures and design patterns for the web
\end{itemize}

\end{frame}

%%%%%%%%%%%%%%%
\begin{frame}
  \frametitle{Additional knowledge:}
Complements to this course:
\begin{itemize}
\item EITF05 - Web Security
\item EITN41 - Advanced Web Security
\item MAMN25 - Interaction Design
\item EDAP10  - Concurrent Programming
\item EDAF75 - Database Technology
\item EDAN40 - Functional Programming
\item EDAG05 - Agil programvaruutveckling - projekt
\end{itemize}

\end{frame}

%%%%%%%%%%%%%%%
\begin{frame}
  \frametitle{Outline}
Course content:
\begin{itemize}
\item 10 lectures in Swedish
\item 4 labs
\item project
\end{itemize}
\vspace{2mm}
All activities are in zoom. You find the links in the schedule.
\\ \vspace{2mm}
Labs must be reported in groups of 2:
\begin{itemize}
 \item solve the lab together
    \begin{enumerate}
      \item collaborate using zoom and/or git.
   \end{enumerate}
 \item solve the assignment independently
   \begin{enumerate}
     \item compare solutions
     \item alternate who's solution you show
   \end{enumerate}
\end{itemize}
\end{frame}

%%%%%%%%%%%%%%%
\begin{frame}
  \frametitle{Outline}
Course content:
\begin{itemize}
\item 10 lectures in Swedish
\item 4 labs
\item project
\end{itemize}
\vspace{2mm}
All activities are in zoom. You find the links in the schedule.
\\ \vspace{2mm}
Labs must be reported in groups of 2:
\begin{itemize}
 \item solve the lab together
    \begin{enumerate}
      \item collaborate using zoom and/or git.
   \end{enumerate}
 \item solve the assignment independently
   \begin{enumerate}
     \item compare solutions
     \item alternate who's solution you show
   \end{enumerate}
\end{itemize}
\end{frame}

%%%%%%%%%%%%%%%
\begin{frame}
  \frametitle{Registration}
To claim your participation:
\begin{itemize}
  \item register in LADOK
  \item enroll for the labs in \href{https://sam.cs.lth.se/LabsSelectSession?occasionId=693}{SAM}
\end{itemize}

\end{frame}

%%%%%%%%%%%%%%%
\begin{frame}
  \frametitle{Study materials}

The course content is best described in several different sources, see the reading instructions.
\\ \vspace{3mm}
Books
\begin{itemize}
\item \href{https://github.com/getify/You-Dont-Know-JS}{You Dont Know JS by Kyle Simpson} 
\item \href{https://eloquentjavascript.net/}{Eloquent JavaScript by Marijn Haverbeke}
\item \href{http://shop.oreilly.com/product/9780596517748.do}{JavaScript: The Good Parts by Douglas Crockford}
\end{itemize}

Other sources
\begin{itemize}
\item \href{http://fileadmin.cs.lth.se/cs/Education/EDAF90/lectures/lecture-notes.pdf}{Lecture notes for EDAF90}
\item \href{https://developer.mozilla.org/en-US/docs/Learn}{Mozilla online material: Learn web development}
\item links to online material on the home page
\end{itemize}

\end{frame}

%%%%%%%%%%%% End frame
\begin{frame}[plain]
  \endpage
\end{frame}

%%%%%%%%%%%
\end{document}
